Parity-odd Neutrino Torque Detection

### 2018-10-25 neu_torque.tex
revised discussion and conclusion

We construct an equivalent inertia $\I_r$ from $\T_c$, in order that $\bs{\epsilon}\,\I_r\T_\nu$ is maximumly correlated with $\spin^I_{\nu 0}$ from Eq.(\ref{eq.jnu0}).
In the primary coordinate of $\T_c$, $\T_c$ can be eigen-decomposed as $\T_c=\sum_{i=1}^3\T^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
\begin{equation}\label{eq.Ir}
	\I_r=\sum_{i=1}^3\alpha_i\T^{\lambda_i}_c
\end{equation}
optimizes the cross-correlation $\mu(\spin^I_{\nu 0},\spin^{I_r})$, where
\begin{equation}\label{eq.jnuIr}
 	\spin^{I_r}_\nu=\bs{\epsilon}\,\I_r\T_\nu
\end{equation}
is the reconstructed neutrino spin direction.

@@

\tcb{\textit{Discussion.---}} 
Accuracy and mass dependency: we notice that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}. We investigate that numerical errors that may affect the results. P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation. Higher mass halos in a given simulation generally have slightly higher $\mu_L$ and $\mu_I$, however the correlation is enhanced greatly as we use higher mass resolutions. All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation. With different box sizes, mass resolutions, force resolutions, and find that the results are consistent across these simulations. Especially, in the estimation of number of halo spins needed to detect the neutrino torque, the deciding number is the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$. This has a very weak dependence on halo mass and configuration of the simulation. An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles or fluids to calculate more precise nonlinear neutrino effects. These require a simulation comparable to the computing resources of TianNu \citep{2017RAA....17...85E}.

Observability: surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P} can cover 1 billion galaxies. There are xxx HI galaxies with redshift 1 \citep{2004MNRAS.350.1210Z}.

Beyond standard models: it was also suggested that neutrino mass is generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}.

Codis \citep{2018MNRAS.tmp.2447C}.

\tcb{\textit{Conclusion.---}} 
Enormous efforts has been contributed to the neutrino mass properties.
Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physic mysteries. 
The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
and is easily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
It contains comparable amount of information as density field but poorly appreciated in application.
The recently developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space.

>>>>

We construct an equivalent inertia $\I_r$ from $\T_c$, in order that $\bs{\epsilon}\,\I_r\T_\nu$ is maximumly correlated with $\spin^I_{\nu 0}$ from Eq.(\ref{eq.jnu0}).
In the primary coordinate of $\T_c$, $\T_c$ can be eigen-decomposed as $\T_c=\sum_{i=1}^3\T^{\lambda_i}_c$ and find $\alpha_i$ such that the reconstructed
\begin{equation}\label{eq.Ir}
	\I_r=\sum_{i=1}^3\alpha_i\T^{\lambda_i}_c
\end{equation}
optimizes the cross-correlation between $\spin^I_{\nu 0}$ and $\spin^{I_r}_\nu$, where
\begin{equation}\label{eq.jnuIr}
 	\spin^{I_r}_\nu=\bs{\epsilon}\,\I_r\T_\nu
\end{equation}
is the reconstructed neutrino spin direction.

@@

\tcb{\textit{Discussion.---}} 
We notice that the cross-correlation coefficients between $\spin_0$ and initial values $\spin_L$ and $\spin_T$ are prominently higher than what was measured in \citep{2000ApJ...532L...5L}. We investigate that numerical errors that may affect the results. P3M (particle-particle particle-mesh) algorithms result in higher cross-correlation between initial and final spins, compared to PM (particle-mesh), where additional tangential forces in PM violate the angular momentum conservation. Higher mass halos in a given simulation generally have slightly higher cross-correlations between initial and final spins, however the correlation is enhanced greatly as we use higher mass resolutions. All other cross-correlation measurements have only weak dependencies on the halo mass, even in a fixed simulation. With different box sizes, mass resolutions, force resolutions, and find that the results are consistent across these simulations. Especially, in the estimation of number of halo spins needed to detect the neutrino torque, the deciding number is the cross-correlation $\left\langle\spin^I_{\nu 0}\cdot\spin^{I_r}_\nu\right\rangle=0.19$. This has a very weak dependence on halo mass and configuration of the simulation. An accurate study of the neutrino torque requires mass and force resolutions to cover the wide range of halo mass, a large box size ($>600\,{\rm Mpc}/h$) to account the neutrino tides at distance, and neutrino particles or fluids to calculate more precise nonlinear neutrino effects. These require a simulation comparable to the computing resources of TianNu \citep{2017RAA....17...85E}.

For surveys like the Hubble Sphere Hydrogen Survey (HSHS) \citep{2006astro.ph..6104P}, we are able to observe \tcr{$10^9$} HI galaxies ($>10^{12}M_\odot$), 
in a cosmic volume $(4{\rm Gpc}/h)^3$ below redshift $z\simeq 1$   \citep{2004MNRAS.350.1210Z}. 
Under the standard model of the Universe with $M_\nu=0.05$ eV, a $5\sigma$ detection will be reachable.
Beyond standard models, e.g., the neutrino mass could be generated by a gravitational $\theta$-term condensation as the source of dark energy \citep{2016PhRvD..93k3002D}, the tidal torque history will be different and the numbers of galaxies are different to differentiate between these models.

%Codis \citep{2018MNRAS.tmp.2447C}.

\tcb{\textit{Conclusion.---}} 
Enormous efforts has been contributed to the neutrino mass properties.
Radio astronomic and cosmological surveys are promising low-cost experiments implemented in the Universe to probe basic physic mysteries. 
The angular momentum is a 3-dimensional, gravity-driven, parity-odd measure, 
and is easily observable, well modeled by the tidal torque theory, and well conserved over the cosmic evolution.
It contains comparable amount of information as density field but poorly appreciated in application.
The recently developed reconstruction techniques enable us to visit an-order-of-magnitude more precise initial conditions of the Universe in Lagrangian space, 
and thus to accurately reconstruct the neutrino torque effects presented in the letter.
